Sparse Function-space Representation \\ of Neural Networks

### 2023-05-19 paper/main.tex
first draft multi-class notation for cl

\newcommand{\GP}{\mathcal{GP}}
\newcommand{\KL}[2]{\mathrm{D}_\textrm{KL} \dbar*{#1}{#2}}
\newcommand{\MKzz}{\mbf{K}_{\mbf{z}\mbf{z}}}
\newcommand{\MKxx}{\mbf{K}_{\mbf{x}\mbf{x}}}
\newcommand{\MKzx}{\mbf{K}_{\mbf{z}\mbf{x}}}
\newcommand{\MKxz}{\mbf{K}_{\mbf{x}\mbf{z}}}
\newcommand{\vkzi}{\mbf{k}_{\mbf{z}i}}
\newcommand{\vkzs}{\mbf{k}_{\mbf{z}i}}
\newcommand{\vk}{\mbf{k}}
\newcommand{\MLambda}[0]{\mathbold{\Lambda}}
\newcommand{\MSigma}[0]{\mathbold{\Sigma}}
\definecolor{matplotlib-blue}{HTML}{1f77b4}
\newcommand{\N}{\mathrm{N}}
%\newcommand{\R}{\mathrm{R}}
\newcommand{\myexpect}{\mathbb{E}}

@@

\section{\our: Sparse function-space representation of neural networks}
%\paragraph{GP in the dual parameters}
The seminal work by \citet{csato2002sparse} (parameterization Lemma~1) gives a parameterization for the posterior process that can be found through the Bayesian update using the GP prior and the likelihood function. This gives a `dual' parameterization, $\valpha$ and $\vbeta$,
%
\begin{equation}  \label{eq:gp_pred}
  \myexpect_{p(f_i \mid\vy)}[f_i]= \vk_{\vx i}^\top \valpha \quad \text{and} \quad
  \mathrm{Var}_{p(f_i \mid \vy)}[f_i] = k_{ii} - \vk_{\vx i}^\top ( \MKxx + \diag(\vbeta)^{-1})^{-1} \vk_{\vx i},
\end{equation}
%
where the $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{N \times N}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, and $k_{ii} = \kappa(\vx_i, \vx_i)$.
\cref{eq:gp_pred} states that the first two moments of the resultant posterior process (which may not be a GP), can be parameterized via the dual
parameters $\valpha, \vbeta \in \R^{N}$,
% which are the vectors
defined as, \looseness-1
%
\begin{equation}
  \label{eq:dual_param}
  \alpha_i \coloneqq \myexpect_{p(\vw \mid \vy)}[\nabla_{f}\log p(y_i \mid f) |_{f=f_i}]
  \quad \text{and} \quad
  \beta_i \coloneqq - \myexpect_{p(\vw \mid \vy)}[\nabla^2_{f f}\log p(y_i \mid f_i) |_{f=f_i}],
\end{equation}
%
where $f_\vw(\vx_i) = f_i$ is the function output at input $\vx_i$. The relationships specified are valid for generic likelihoods, and involve no approximations since the expectation is under the exact posterior, but given that the model can be expressed in a kernel formulation. \cref{eq:gp_pred} and \cref{eq:dual_param} highlight that the approximate inference technique, usually viewed as a posterior approximation, can be alternatively interpreted as an approximation of the expectation of gradients of loss/likelihoods.

@@

\subsection{Extending the CL regularizer to multi-class settings}
\label{sec:cl_multioutput} \todo{this sub-section was mistakenly duplicated}
\begin{equation}
	\bar{\MB}^{-1}_s = \MKzz^{-1} \vbeta_\vu \MKzz^{-1} \in \R^{C \times m \times m} \quad \MKzz \in \R^{C \times m \times m} 
\end{equation}

\begin{equation}
	\mathcal{R_\textit{SFR}}(\mathbf{w}) = \sum_{s=1}^{t-1}	\sum_{k \in 	C}\left[\left(f_{\vw, k}(\MZ_{s}) - f_{\vw_{s}, k}(\MZ_s) \right)^\T \bar{\MB}^{-1}_{s} \left(f_{\vw, k}(\MZ_{s}) - f_{\vw_{s, k}}(\MZ_s) \right) \right] 
\end{equation}

>>>>

\newcommand{\GP}{\mathcal{GP}}
\newcommand{\KL}[2]{\mathrm{D}_\textrm{KL} \dbar*{#1}{#2}}
\newcommand{\MKzz}{\mbf{K}_{\mbf{z}\mbf{z}}}
\newcommand{\MKzzc}{\mbf{K}_{\mbf{z}\mbf{z}, c}}
\newcommand{\MKxx}{\mbf{K}_{\mbf{x}\mbf{x}}}
\newcommand{\MKzx}{\mbf{K}_{\mbf{z}\mbf{x}}}
\newcommand{\MKxz}{\mbf{K}_{\mbf{x}\mbf{z}}}
\newcommand{\vkzi}{\mbf{k}_{\mbf{z}i}}
\newcommand{\vkzic}{\mbf{k}_{\mbf{z}i,c}}
\newcommand{\vkzs}{\mbf{k}_{\mbf{z}i}}
\newcommand{\vk}{\mbf{k}}
\newcommand{\MLambda}[0]{\mathbold{\Lambda}}
\newcommand{\MSigma}[0]{\mathbold{\Sigma}}
\definecolor{matplotlib-blue}{HTML}{1f77b4}
\newcommand{\N}{\mathrm{N}}
%\newcommand{\R}{\mathrm{R}}
\newcommand{\myexpect}{\mathbb{E}}

@@

\section{\our: Sparse function-space representation of neural networks}\label{sec:sfr}
%\paragraph{GP in the dual parameters}
The seminal work by \citet{csato2002sparse} (parameterization Lemma~1) gives a parameterization for the posterior process that can be found through the Bayesian update using the GP prior and the likelihood function. This gives a `dual' parameterization, $\valpha$ and $\vbeta$,
%
\begin{equation}  \label{eq:gp_pred}
  \myexpect_{p(f_i \mid\vy)}[f_i]= \vk_{\vx i}^\top \valpha \quad \text{and} \quad
  \mathrm{Var}_{p(f_i \mid \vy)}[f_i] = k_{ii} - \vk_{\vx i}^\top ( \MKxx + \diag(\vbeta)^{-1})^{-1} \vk_{\vx i},
\end{equation}
%
where the $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{N \times N}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, and $k_{ii} = \kappa(\vx_i, \vx_i)$.
\cref{eq:gp_pred} states that the first two moments of the resultant posterior process (which may not be a GP), can be parameterized via the dual
parameters $\valpha, \vbeta \in \R^{N}$,
% which are the vectors
defined as, \looseness-1
%
\begin{equation}
  \label{eq:dual_param}
  \alpha_i \coloneqq \myexpect_{p(\vw \mid \vy)}[\nabla_{f}\log p(y_i \mid f) |_{f=f_i}]
  \quad \text{and} \quad
  \beta_i \coloneqq - \myexpect_{p(\vw \mid \vy)}[\nabla^2_{f f}\log p(y_i \mid f_i) |_{f=f_i}],
\end{equation}
%
where $f_\vw(\vx_i) = f_i$ is the function output at input $\vx_i$. The relationships specified are valid for generic likelihoods, and involve no approximations since the expectation is under the exact posterior, but given that the model can be expressed in a kernel formulation. \cref{eq:gp_pred} and \cref{eq:dual_param} highlight that the approximate inference technique, usually viewed as a posterior approximation, can be alternatively interpreted as an approximation of the expectation of gradients of loss/likelihoods.

@@

\subsection{Extending the CL regularizer to multi-class settings} 
\label{sec:cl_multioutput}
In the multi-class setting, the neural network can be seen as a function approximator that maps inputs $\vx \in \inputDomain$ to  the logits $\textit{\textbf{f}} \in \outputDomain$, \ie, $f_\mathbf{w}: \inputDomain \to \outputDomain$. As a consequence, the Jacobian matrix is $\Jac{\weights}{\vx} \coloneqq \left[ \nabla_\weights f_\weights(\vx)\right]^\top \in \R^{C \times P}$ and the kernel of the associated sparse GP derived in \cref{sec:sfr}, $\kappa(\vx, \vx')$ is a tensor in $\R^{C \times P \times P}$. This means that we need compute the dual parameters and the matrix $\bar{\MB}^{-1}_{t, c}$ for all the function's outputs. If we denote $\kappa_c(\vx, \vx') \in \R$, as the kernel related to the $c$\textsuperscript{th} output function, \ie, the logit for class $c$, we can write the dual parameters as
%
\begin{equation} \textstyle
  \valpha_{\vu, c}  =  \sum_{i=1}^N  \vkzic \, \hat{\alpha}_{i, c} \in \R^{M}
  \quad \text{and} \quad
  \MBeta_{\vu, c} =  \sum_{i=1}^N \vkzic \,\hat{\beta}_{i, c} \, \vkzic^{\T} \in \R^{M \times M} ,    
\label{eq:dual_sparse}
\end{equation}
%
where $\vkzic$ is the kernel $\kappa_c(\cdot, \cdot)$ computed for the inducing points $\vz \in \mathcal{Z}$ and the input point $\vx_i$ for class $c$. In this setting, the dual variables are
\begin{equation}
\begin{aligned}
  \hat{\alpha}_{i,c} &\coloneqq  \left[\hat{\valpha}_{i}\right]_c = \left[\nabla_{f}\log p(y_i \mid f) |_{f=f_i} \right]_c \in \R\\
  \hat{\beta}_{i,c} &\coloneqq \left[\textrm{diag}(\hat{\vbeta}_i)\right]_c = \left[ - \nabla^2_{ff}\log p(y_i \mid f) |_{f=f_i} \right]_c \in \R 
\end{aligned}
\end{equation}
where the operator $[ \cdot ]_c$ is taking the $c$\textsuperscript{th} element of a vector. In fact, in the multi-class setting $\hat{\valpha}_{i}$ is a vector in $\R^C$ and $\hat{\vbeta}_{i}$ a matrix in $\R^{C \times C}$.
For the \our-regularizer, at the end of each task we compute the matrix $\bar{\MB}^{-1}_{t, c}$ as 
\begin{equation}\textstyle
 	\bar{\MB}^{-1}_{t, c} = \MKzzc^{-1} \MBeta_{\vu, c} \MKzzc^{-1} \in \R^{M \times M}, 
 	\quad \text{s.t.} \quad
 	\MBeta_{\vu, c} =  \sum_{i \in \dataset_t} \vkzi \,\hat{\beta}_{i, c} \, \vkzi^{\T}.
\end{equation}
However, in practice computing the kernel $\kappa_c$ for classes $c$, for which the model has not seen any data yet is not meaningful \todo{probably better to word this differently}, so we define $\bar{\MB}^{-1}_{t, c}$ to be  $\bar{\MB}^{-1}_{t, c} = \MI_M$ $\forall c \not\in \mathcal{C}_t$, where $\mathcal{C}_t$ is the set of classes observed up to task t. For example, for S-MNIST  we will have after task 1 $\mathcal{C}_1 = \{0, 1\}$, then after task 2 $\mathcal{C}_1 = \{0, 1, 2, 3\}$, and so on. 
Finally, we can rewrite $\mathcal{R}_\mathrm{\our}$ as follows
\begin{equation}\label{eq-cl-regularizer}\textstyle
\begin{aligned}
  \mathcal{R}_\mathrm{\our}(\weights, \mathcal{M}) = \frac{1}{2} \sum_{s=1}^{t-1} \sum_{c=1}^{C} \frac{1}{M} 
	\left\lVert 
	\vu_c - \vu_{s,c} % f_{\weights_{s}}(\MZ_s)
	\right\rVert_{\bar{\MB}^{-1}_{s}}, \\
	\text{~with~} \vu_c \in \R^M \text{~with~} u_{i, c} = \left[ f_{\weights}(\vz_{i}) \right]_c,~\forall {\vz_i} \in \mathcal{Z}_s .
\end{aligned}
\end{equation}
